---
title: "Retail Analysis"
author: "StepUp Analytics"
date: "02/06/2021"
output: word_document
---

```{r message=F, warning=FALSE}
install.packages('Hmisc')
install.packages('dplyr')
install.packages('ggplot2')

library(dplyr)
library(ggplot2)
library(Hmisc)
```

*Setting and getting working directory*

```{r}
path = '/Users/md.zishan/Documents/S/StepUp Analytics/Learn and Intern/R Programming/Assignment'
setwd(path)
getwd()
```

*Reading the files*

```{r}
customer_full_df <- read.csv(
  'Customer.csv',
  header = TRUE,
  stringsAsFactors = FALSE,
  na.strings = c("", "NA")
)

transactions_full_df <- read.csv(
  'Transactions.csv',
  header = TRUE,
  stringsAsFactors = FALSE,
  na.strings = c("", "NA")
)

prodcat_full_df <- read.csv(
  'prod_cat_info.csv',
  header = TRUE,
  stringsAsFactors = FALSE,
  na.strings = c("", "NA")
)

```

**1. Merge the datasets Customers, Product Hierarchy and Transactions as Customer_Final. Ensure to keep all customers who have done transactions with us and select the join type accordingly.**

**1. a) Merge using the base merge()**

```{r}
trans_cust_df1 <- merge(
  transactions_full_df,
  customer_full_df,
  by.x = "cust_id",
  by.y = "customer_Id",
  all.x = TRUE,
  all.y = FALSE
)

cust_fin1 <- merge(
  trans_cust_df1,
  prodcat_full_df,
  by.x = c("prod_cat_code", "prod_subcat_code"),
  by.y = c("prod_cat_code", "prod_sub_cat_code"),
  all.x = TRUE,
  all.y = FALSE
)

cat('Dimensions of dataframe obtained through merge(): ', dim(cust_fin1))
```

*1. b) Merge using Dplyr merge functions*

```{r}
trans_cust_df2 <- left_join(transactions_full_df,
                            customer_full_df,
                            by = c("cust_id" = "customer_Id"))

Customer_Final <- left_join(
  trans_cust_df1,
  prodcat_full_df,
  by = c("prod_cat_code" = "prod_cat_code",
         "prod_subcat_code" = "prod_sub_cat_code")
)

cat('Dimensions of dataframe obtained through dplyr(): ', dim(Customer_Final))
```

**2. Prepare a summary report for the merged data set.**

*2. a) Get the column names and their corresponding data types.*

```{r}

summary(Customer_Final)
str(Customer_Final$Qty)
```

*2. b) Top/Bottom 10 observations*

```{r}
head(Customer_Final,10)
tail(Customer_Final,10)
```

*2. c) “Five-number summary” for continuous variables (min, Q1, median,Q3 and max)*

```{r}
summary(Customer_Final$Tax)

```

*2. d) Frequency tables for all the categorical variables*

```{r}
factor(Customer_Final$Store_type)
```

**3. Generate histograms for all continuous variables and frequency bars for categorical variables.**

*Histograms for continuous variables*

```{r}
h1 = ggplot(Customer_Final, aes(x = Qty)) +
  geom_histogram(
    # binwidth = 0.1,
    color = "red",
    linetype = "dashed",
    fill = "black"
  )

plot(h1)

```

*Frequency bars for continuous variables*

```{r}
b1 = ggplot(Customer_Final, aes(x = Qty)) +
  geom_boxplot(
  
    color = "Yellow",
    linetype = "dashed",
    fill = "black"
  )

plot(b1)
```

**4. Calculate the following information using the merged dataset :**

*4 a) Calculate Time period of the available transaction data.*

```{r}

```

*4 b) Calculate count of transactions where the total amount of transaction was negative*

```{r}

a = Customer_Final %>% count(total_amt < 0)
rownames(a) = c('Positive','Negative')
colnames(a) = c('Transactions','total_amt')
a

```

**5. Analyze which product categories are more popular among females vs male customers.**

*Most popular categories among males*

```{r}

Customer_Final
male = group_by(Customer_Final,prod_cat)

summarise(
  male,
  only_male = sum(Gender=="M",na.rm = TRUE) 
)





```

*Most popular categories among Females*

```{r}
Customer_Final
female = group_by(Customer_Final,prod_cat)
summarise(
  female,
  only_Female = sum(Gender=="F", na.rm = TRUE)
  
)


```

*The top 3 categories for both males and females are:*

* Books

* Electronics

* Home and kitchen

**6. Which City code has the maximum customers and what was the percentage of customers from that city?**

*Percentage of customers belonging to different city codes sorted in descending order:*

```{r}
Customer_Final
a6 = arrange(Customer_Final,city_code)
a6




```

The highest percentage of customers come from city code 4 (10.51%)

**7. Which store type sells the maximum products by value and by quantity?**

*Products sold by value:*

```{r}
Customer_Final
stores_grp = group_by(Customer_Final,Store_type)
summarise(
  stores_grp,
  sold_by_value = sum(total_amt) )


```

```{r}
Customer_Final
stores_grp = group_by(Customer_Final,Store_type)
summarise(
  stores_grp,
  sold_by_Quantity = sum(Qty) )

```

**8. What was the total amount earned from the "Electronics" and "Clothing"categories from Flagship Stores?**

```{r}
Customer_Final
a8 = group_by(Customer_Final,Store_type, prod_cat)

summarise(a8,sum(total_amt))



```

**9. What was the total amount earned from "Male" customers under the "Electronics" category?**

```{r}
# Customer_Final
# a9 = group_by(Customer_Final,Gender,prod_cat)
# 
# summarise(a9,sum(total_amt,na.rm = TRUE))
Customer_Final
a9 = group_by(Customer_Final,prod_cat)

summarise(
  a9,
  only_male = sum(Gender=="M",na.rm = TRUE) 
)


```

**10. How many customers have more than 10 unique transactions, after removing all transactions which have any negative amounts?**

```{r}
Customer_Final
a10 = filter(Customer_Final,total_amt>0)
a10
summarise(
  a10,
  after_removing_negative = sum(total_amt))


```